Long term cloud enhancement events of global solar irradiance over Thessaloniki, Greece

### 2024-06-20 article/article.tex
Commit 2024-06-20 10:15

We studied the enhancement of the global horizontal irradiance (GHI),
over Thessaloniki, for the period 1993 -- 2023, on one minute average
records. We identify the cloud enhancement events (CE), by creating an
appropriate clear sky irradiance reference with the use of a radiation
transfer model, and aerosol optical depth (AOD) data, from the AERONET,
and a collocated Brewer photospectrometer. We found that there is a
trend of CE cases of \(45.6\pm 21.9\,\text{cases}/\text{year}\), and a
mean total energy change, of the CE events by
\(116.9\pm 67.8\,\text{kJ}/\text{year}\) with the peak of the CE events
are observed during May and June. An analysis of the total duration of
CE events, showed that durations longer than 5 minutes are very rare,
with exceptions lasting over an hour. We have detected enhancements over
the total solar irradiance, without exceeding the maximum values
reported for other sites, with more favorable CE conditions.

@@

An important aspect of the variability of solar radiation is its
interaction with the clouds. In general, clouds can attenuate a fraction
of solar irradiance, but under certain conditions, can lead to
enhancement of the global horizontal irradiance (GHI) reaching the
ground. This effect can locally increase the observed GHI to levels even
higher than the expected clear-sky irradiance \citep[ and references
therein]{Cordero2023, Vamvakas2020, CastillejoCuberos2020, Vamvakas2020}.

Some of the proposed underling mechanisms of those enhancements, have
been summarized by \citet{Gueymard2017}; the most important being the
scattering of radiation on the edges of cumulus clouds. It has been also
suggested that enhancement of GHI can be produced by thin cirrus clouds
though refraction and scattering \citep{Thuillier2013}. Further
investigation with radiative transfer modeling and observations pointed
as the prevailing mechanism, the strong forward Mie scattering through
clouds of low optical depth
\citep{Pecenak2016, Thuillier2013, Yordanov2013, Yordanov2015}. Overall,
enhancement event depend on different interactive factors, which include
cloud thickness, structure and type, and the relative position of the
sun and the clouds \citep{Gueymard2017, Veerman2022}.

On multiple sites, cloud enhancements (CE) have been reported to be able
to briefly exceed the value of the solar constant, resulting of
clearness indices above unit. A summary of extreme enchantment cases
(ECE) has been compiled by \citet{Martins2022}. There are also some
practical implications of the cloud enchantments. The intensity and
duration of enhancements can effect the efficiency and stability of
photovoltaic power production \citep{Lappalainen2020, Jaervelae2020}
while ECEs have the potential to compromise the integrity of
photovoltaic plants infrastructure \citep{DoNascimento2019}. It has also
been demonstrated that these events can interfere in the comparison of
ground-based and satellite observations \citep{Damiani2018}. Global
warming has likely affected cloud coverage in the last few decades.
\citet{Liu2023} reported increases in cloud cover over the tropical and
subtropical oceans and decreases over most continents, while
\citet{Dong2023} reported decreases over North America and Europe. To
our knowledge, there is no evidence on whether this trend has affected
also the number and strength of CE events.

>>>>

We studied the enhancement of the global horizontal irradiance induced
by clouds over Thessaloniki for the period 1993 -- 2023 using data
recorded every one minute. We identified the cloud enhancement events
(CE) by creating an appropriate clear sky irradiance reference with the
use of a radiative transfer model and aerosol optical depth data from a
collocated Cimel sun photometer and a Brewer spectrophotometer. We found
a trend in CE events of \(45.6\pm 21.9\,\text{cases}/\text{year}\), and
a trend in the CE events irradiation of
\(116.9\pm 67.8\,\text{kJ}/\text{year}\). The peak of the CE events was
observed during May and June. An analysis of the total duration of CE
events, showed that durations longer than 5 minutes are very rare, with
exceptions lasting over an hour. Finally we have detected enhancements
above the total solar irradiance at the top of the atmosphere of up to
\(400\,\text{W}/\text{m}^{2}\), with the \(75\,\%\) of the cases below
\(200\,\text{W}/\text{m}^{2}\).

@@

An important aspect of the variability of solar radiation is its
interaction with the clouds. In general, clouds can attenuate a fraction
of solar irradiance, but under certain conditions, can lead to
enhancement of the global horizontal irradiance (GHI) reaching the
ground. This cloud enhancement (CE) effect can locally increase the
observed GHI to levels even higher than the expected clear-sky
irradiance \citep[ and references
therein]{Cordero2023, Vamvakas2020, CastillejoCuberos2020, Vamvakas2020}.

Some of the proposed underling mechanisms of those enhancements, have
been summarized by \citet{Gueymard2017}; the most important being the
scattering of radiation on the edges of cumulus clouds. It has been also
suggested that enhancement of GHI can be produced by thin cirrus clouds
though refraction and scattering \citep{Thuillier2013}. Further
investigation with radiative transfer modeling and observations pointed
as the prevailing mechanism, the strong forward Mie scattering through
clouds of low optical depth
\citep{Pecenak2016, Thuillier2013, Yordanov2013, Yordanov2015}. Overall,
enhancement events depend on different interactive factors, which
include cloud thickness, structure and type, and the relative position
of the sun and the clouds \citep{Gueymard2017, Veerman2022}.

On multiple sites, cloud enhancements have been reported that exceed
momentarily the solar constant, resulting in clearness indices above
unit. A summary of extreme cloud enchantment (ECE) cases has been
compiled by \citet{Martins2022}. Cloud enhancements can have also some
practical implications. The intensity and duration of enhancements can
effect the efficiency and stability of photovoltaic power production
\citep{Lappalainen2020, Jaervelae2020} while ECEs have the potential to
compromise the integrity of photovoltaic plants infrastructure
\citep{DoNascimento2019}. It has also been demonstrated that these
events can interfere in the comparison of ground-based and satellite
observations \citep{Damiani2018}. Global warming has likely affected
cloud coverage in the last few decades. \citet{Liu2023} reported
increases in cloud cover over the tropical and subtropical oceans and
decreases over most continents, while \citet{Dong2023} reported
decreases over North America and Europe. To our knowledge, there is no
evidence on whether this trend has affected also the number and strength
of CE events.
